
```{r, echo=FALSE}
rm(list=ls())

## libraries
suppressPackageStartupMessages(library(network))
suppressPackageStartupMessages(library(ergm))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(reshape2))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(data.table))
suppressPackageStartupMessages(library(plyr))
```


```{r, echo=FALSE}

location <- "../../Release/out_expID2"
#sim_num <- 
dir_nums <- 1:5
ntimesteps <- 25000

dirs <- lapply(dir_nums, function (i) paste0(location, "/output_", i))

## read data
   ## networks
   main_nets <- lapply(dirs, function (i) 
                        readRDS(paste0(i, "/main_network_", ntimesteps, ".RDS"))
                       ) #main

   cas_nets <- lapply(dirs, function (i) 
                        readRDS(paste0(i, "/casual_network_", ntimesteps, ".RDS"))
                       )

   biom_data <- lapply(dirs, function (i) 
                             fread(paste0(i, "/biomarker_log.csv"))
                       )

   counts_data <- lapply(dirs, function (i) 
                               fread(paste0(i, "/counts.csv"))
                         )

   inf_event_data <- lapply(dirs, function (i) 
                                  fread(paste0(i, "/infection_events.csv"))
                            )

   pt_event_data <- lapply(dirs, function (i) 
                                  fread(paste0(i, "/partnership_events.csv"))
                            )
   

   person_data <- lapply(dirs, function (i)
                               fread(paste0(i, "/person_data.csv"))
                         )

   ## parameters
   source("../common/chi_params_nonderived.R")
   source("../common/chi_params_derived.R")
```

These datasets include the sexual network at the last time step, biomarker data with detailed trajectories, infection, death and partnership events, and counts of various quantities at each time step.

# Results
## Demography
We first compute the annual growth rate.  

```{r, echo=FALSE, results='hide'}
   
  final.vcount <- unlist(lapply(main_nets, network.size))
  init.vcount <- 5000
  
  annual.growth.rate <- (((final.vcount)/init.vcount)^(1/(10e3/365))-1)*100 #log scale
```

The average final population size, across the `r length(dir_nums)` is 
`r mean(final.vcount)`, corresponding to a growth rate of 
`r round(mean(annual.growth.rate), 2)`%. (The rate of entry of new individuals is a free parameter, and will be adjusted to reflect data form the three cities). 

The age distribution at the end of the simulation is below.


```{r, results='show'}
# gg net%v%"age"
# plot(
#   qplot(age, geom="histogram", binwidth=5, 
#        col=I("red"), 
#        alpha=I(.2))
#   )
```

## Prevalence
    
   ```{r, echo=FALSE, results='hide'}

   alive_person_data <- lapply(person_data, function (i)
                               i[`time of death` == -1]
                               )

   final_popsize <- lapply(alive_person_data, nrow)

   infected_data <- lapply(alive_person_data, function(i)
                          i[`infection status` == 1]
                          )
   n_infected_end <- lapply(infected_data, nrow)

   prev <- unlist(n_infected_end)/unlist(final_popsize)
   
   ```
  The overall prevalence at the end of the simulation was about `r round(prev, 1)`%.  HIV prevalence as a function of time is shown below.
  
  ```{r, echo=FALSE}
     counts_data <- lapply(counts_data, function(i)
                                            i[-1,]
                         ) 

     counts_data_df <- do.call("rbind", lapply(counts_data, as.data.frame))
     sim_num <- ((1:nrow(counts_data_df)) %/% (25001))+1
     sim_num <- as.factor(sim_num)
     counts_data_df$sim_num <- sim_num

     ggplot(counts_data_df, 
          aes(x=time/365, 
             y=(vertex_count-uninfected)/vertex_count,
             color=sim_num))+
         geom_line()+
         scale_y_continuous(limits=c(0,1))+
         ylab("Prevalence")+
         xlab("Time (years)")
  ```
   
## Incidence  

The annual incidence rates over the course of the simulation are shown below.

   ```{r, results='hide', echo=FALSE}
  
      inc_per_timestep <- lapply(counts_data, function (i)
                                           (i$infected_via_transmission)/
                                           (i$uninfected - i$entries)
                                           )

      summ_inc_per_timestep <- lapply(inc_per_timestep, summary)
     
      inc_per_year <- lapply(inc_per_timestep, 
                                         function(i)
                                         split(i, 
                                         ceiling(seq_along(i)/365))
                                         )

      list_mean_inc_per_year <- lapply(inc_per_year, function (i) lapply(i, mean))
      summ_inc_across_sim <- summary(unlist(list_mean_inc_per_year))*365*100

      list_mean_inc_per_year_df <-   do.call("rbind", lapply(list_mean_inc_per_year, 
                                            as.data.frame))
      colnames(list_mean_inc_per_year_df) <- as.numeric(1:69)
      list_mean_inc_per_year_df <- melt(list_mean_inc_per_year_df)
      list_mean_inc_per_year_df$sim_num <- as.factor(1:5)

      list_mean_inc_per_year_df[] <- lapply(list_mean_inc_per_year_df, as.numeric)
      
      ggplot(list_mean_inc_per_year_df, 
             aes(x=variable, y=value*365*100, color=as.factor(sim_num)))+geom_line()
     
   ```
   
### Main Partnerships

#### Mean degree. 
The target main mean degree is `r round(nedges*2/n, 2)` and a plot of the simulated mean
degree, over the course of the simulations, is below.

```{r, echo=FALSE, show='hide'}
   #counts_data <- rbind(counts_data, nedges)
   #type <- c(rep("simulated", (nrow(counts_data)-1)), rep("target", 1))
   #counts_data <- as.data.frame(cbind(counts_data, type))
   
  ggplot(counts_data_df)+
        geom_line(aes(x=time/365, y=main_edge_count*2/vertex_count, col=sim_num))+
        geom_line(aes(x=time/365, y=nedges*2/n))+
        xlab("time step")+ylab("mean degree")+
        scale_y_continuous(limits=c(0.4, 0.5))+
        ylab("Mean degree (main)")+
        xlab("Time (years)")
  
```

#### Degree Distribution

```{r, echo=FALSE, show='hide'}
   sim_deg_dist_in_num <- lapply(main_nets, degreedist)
   sim_deg_dist_prop <- unlist(sim_deg_dist_in_num)/unlist(final.vcount)
   sim_deg_dist_prop_mat <- matrix(sim_deg_dist_prop, ncol=4, byrow=TRUE)

   sim_deg_dist_prop_mat_mean <- apply(sim_deg_dist_prop_mat, 2, mean)
   names(sim_deg_dist_prop_mat_mean) <- as.factor(0:3)
   
   target_deg_dist_in_num <- c(deg_seq, 0)
   target_deg_dist_prop <- target_deg_dist_in_num/init.vcount
   names(target_deg_dist_prop) <-  names(sim_deg_dist_prop_mat_mean) 
  
   main_deg_dist_data <- cbind(t(sim_deg_dist_prop_mat_mean), t(target_deg_dist_prop))
   main_deg_dist_data <- melt(main_deg_dist_data)
   main_deg_dist_data$type <- c(rep("mean_of_sim", 4), rep("target", 4))
   main_deg_dist_data_df <- as.data.frame(main_deg_dist_data)
   

   ggplot(main_deg_dist_data_df, aes(x=Var2, y=value, fill=type))+
    geom_bar(position="dodge", stat="identity")+
    ylab("Proportion of Agents")+
    xlab("Degree")

```

#### Partnership duration.
```{r, echo=FALSE, results='hide'}
## (fast method from jonathan)

# looking for mean partnership duration
pt_data_main <- lapply(pt_event_data, function (i)
                                      i[network_type == 0]
                       )
                       
# group by p1, p2
pt_main_counts <- lapply(pt_data_main, 
                        function (i)
                        i[,.(counts = .N),by=.(p1,p2),][,unique(counts)] # 1 2 4 3
                        )
pt_main_counts_n <- lapply(pt_data_main, 
                           function (i)
                           i[,.(counts = .N),by=.(p1,p2),][,.N,by=counts]
                           )
# counts     N
# 1:      1  5501
# 2:      2 21833
# 3:      4    10
# 4:      3     1

# so focus on just 2 is justified
    ## (old DONOTUSE: res <- pt_data_main[,.(tick, counts = .N),by=.(p1,p2),][counts == 2,abs(diff(tick))])

res_list <- lapply(pt_data_main, 
                   function (i)
                     i[,.(tick, counts = .N),by=.(p1,p2),][counts == 2,.(dur = abs(diff(tick))), by=.(p1,p2)][,dur]
                   )
summary(unlist(res_list))

res_dt <- as.data.frame(unlist(res_list))
  qplot(res_dt, geom="histogram", binwidth=100, 
       col=I("red"), 
       alpha=I(.2))

sum.res <- summary(unlist(res_list))

```

### Casual Partnerships

#### Mean degree. 
The target main mean degree is `r round(cas_n_edges*2/n, 2)` and a plot of the simulated mean
degree, over the course of the simulations, is below.

```{r, echo=FALSE, show='hide'}
   #counts_data <- rbind(counts_data, nedges)
   #type <- c(rep("simulated", (nrow(counts_data)-1)), rep("target", 1))
   #counts_data <- as.data.frame(cbind(counts_data, type))
   
  ggplot(counts_data_df)+
        geom_line(aes(x=time/365, y=casual_edge_count*2/vertex_count, col=sim_num))+
        geom_line(aes(x=time/365, y=cas_n_edges*2/n))+
        ylab("Mean degree (casual)")+
        xlab("Time (years)")
  
```


```{r, echo=FALSE, show='hide'}
   sim_cas_deg_dist_in_num <- lapply(cas_nets, degreedist)
   sim_cas_deg_dist_prop <- lapply(dir_nums, 
                               function (i) 
                               unlist(sim_cas_deg_dist_in_num[i])/unlist(final.vcount[i]))
   sim_cas_deg_dist_prop_df <- do.call("rbind.fill", 
                                        lapply(sim_cas_deg_dist_prop, as.data.frame))

    
   cas_n_deg <- unlist(lapply(sim_cas_deg_dist_prop, length))
   pholder <- list()
   list_cas_degs <- lapply(dir_nums, function (i) pholder[[i]] <- (0:(cas_n_deg[[i]] - 1)))
   sim_cas_deg_dist_prop_df$degree <- as.character(unlist(list_cas_degs))

   sim_cas_deg_mean <- aggregate(sim_cas_deg_dist_prop_df[,1], 
                                list(sim_cas_deg_dist_prop_df$degree), mean)
   colnames(sim_cas_deg_mean) <- c("degree", "mean")
   
   names(cas_deg_seq) <- as.character(0:(length(cas_deg_seq)-1))
   target_deg_dist_prop <- as.data.frame(cbind(names(cas_deg_seq),
                                         cas_deg_seq/init.vcount)
                                         )
   colnames(target_deg_dist_prop) <- colnames(sim_cas_deg_mean)

   casual_deg_dist_data <- rbind(sim_cas_deg_mean, target_deg_dist_prop)
   casual_deg_dist_data$type <- c(rep("mean_of_sim", nrow(sim_cas_deg_mean)), 
                                      rep("target", nrow(target_deg_dist_prop)))
   #casual_deg_dist_data_df <- as.data.frame(casual_deg_dist_data)
   
   class(casual_deg_dist_data$mean) <- "numeric"

   ggplot(casual_deg_dist_data, aes(x=degree, y=mean, fill=type))+
    geom_bar(position="dodge", stat="identity")+
    ylab("Proportion of Agents")+
    xlab("Degree")

```