
```{r, echo=FALSE}
rm(list=ls())

## libraries
suppressPackageStartupMessages(library(network))
suppressPackageStartupMessages(library(ergm))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(reshape2))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(data.table))

```


```{r, echo=FALSE}

location <- "../../Release/out_expID2"
#sim_num <- 
dir_nums <- 1:5
ntimesteps <- 25000

dirs <- lapply(dir_nums, function (i) paste0(location, "/output_", i))

## read data
   ## networks
   main_nets <- lapply(dirs, function (i) 
                        readRDS(paste0(i, "/main_network_", ntimesteps, ".RDS"))
                       ) #main

   cas_nets <- lapply(dirs, function (i) 
                        readRDS(paste0(i, "/main_network_", ntimesteps, ".RDS"))
                       )

   biom_data <- lapply(dirs, function (i) 
                             fread(paste0(i, "/biomarker_log.csv"))
                       )

   counts_data <- lapply(dirs, function (i) 
                               fread(paste0(i, "/counts.csv"))
                         )

   inf_event_data <- lapply(dirs, function (i) 
                                  fread(paste0(i, "/infection_events.csv"))
                            )

   pt_event_data <- lapply(dirs, function (i) 
                                  fread(paste0(i, "/partnership_events.csv"))
                            )
   

   person_data <- lapply(dirs, function (i)
                               fread(paste0(i, "/person_data.csv"))
                         )

   ## parameters
   source("../common/chi_params_nonderived.R")
   source("../common/chi_params_derived.R")
```

These datasets include the sexual network at the last time step, biomarker data with detailed trajectories, infection, death and partnership events, and counts of various quantities at each time step.

# Results
## Demography
We first compute the annual growth rate.  

```{r, echo=FALSE, results='hide'}
   
  final.vcount <- unlist(lapply(main_nets, network.size))
  init.vcount <- 5000
  
  annual.growth.rate <- (((final.vcount)/init.vcount)^(1/(10e3/365))-1)*100 #log scale
```

The average final population size, across the `r length(dir_nums)` is 
`r mean(final.vcount)`, corresponding to a growth rate of 
`r round(mean(annual.growth.rate), 2)`%. (The rate of entry of new individuals is a free parameter, and will be adjusted to reflect data form the three cities). 

The age distribution at the end of the simulation is below.


```{r, results='show'}
age <- net%v%"age"
plot(
  qplot(age, geom="histogram", binwidth=5, 
       col=I("red"), 
       alpha=I(.2))
  )
```

## Prevalence
    
   ```{r, echo=FALSE, results='hide'}

   alive_person_data <- lapply(person_data, function (i)
                               i[`time of death` == -1]
                               )

   final_popsize <- lapply(alive_person_data, nrow)

   infected_data <- lapply(alive_person_data, function(i)
                          i[`infection status` == 1]
                          )
   n_infected_end <- lapply(infected_data, nrow)

   prev <- unlist(n_infected_end)/unlist(final_popsize)
   
   ```
  The overall prevalence at the end of the simulation was about `r round(prev, 1)`%.  HIV prevalence as a function of time is shown below.
  
  ```{r, echo=FALSE}
     counts_data <- lapply(counts_data, function(i)
                                            i[-1,]
                         ) 

     counts_data_df <- do.call("rbind", lapply(counts_data, as.data.frame))
     sim_num <- ((1:nrow(counts_data_df)) %/% (25001))+1
     sim_num <- as.factor(sim_num)
     counts_data_df$sim_num <- sim_num

     ggplot(counts_data_df, 
          aes(x=time/365, 
             y=(vertex_count-uninfected)/vertex_count,
             color=sim_num))+
         geom_line()+
         scale_y_continuous(limits=c(0,1))+
         ylab("Prevalence")+
         xlab("Time (years)")
  ```
   
## Incidence  

The annual incidence rates over the course of the simulation are shown below.

   ```{r, results='hide', echo=FALSE}
  
      inc_per_timestep <- lapply(counts_data, function (i)
                                           (i$infected_via_transmission)/
                                           (i$uninfected - i$entries)
                                           )

      summ_inc_per_timestep <- lapply(inc_per_timestep, summary)
     
      inc_per_year <- lapply(inc_per_timestep, 
                                         function(i)
                                         split(i, 
                                         ceiling(seq_along(i)/365))
                                         )

      list_mean_inc_per_year <- lapply(inc_per_year, function (i) lapply(i, mean))
      summ_inc_across_sim <- summary(unlist(list_mean_inc_per_year))*365*100

      list_mean_inc_per_year_df <-   do.call("rbind", lapply(list_mean_inc_per_year, 
                                            as.data.frame))
      colnames(list_mean_inc_per_year_df) <- as.numeric(1:69)
      list_mean_inc_per_year_df <- melt(list_mean_inc_per_year_df)
      list_mean_inc_per_year_df$sim_num <- as.factor(1:5)

      list_mean_inc_per_year_df[] <- lapply(list_mean_inc_per_year_df, as.numeric)
      
      ggplot(list_mean_inc_per_year_df, 
             aes(x=variable, y=value*365*100, color=as.factor(sim_num)))+geom_line()
     
   ```

  
  The mean annual incidence, averaged over the length of the simulation, is `r round(summ_inc_per_timestep[4]*365*100, 2)`%.   
  
## Sexual Networks

### Main Partnerships

#### Mean degree. 
The target main mean degree is `r round(nedges*2/n, 2)` and a plot of the simulated mean
degree, over the course of the simulation, is below.

```{r, echo=FALSE, show='hide'}
   #counts_data <- rbind(counts_data, nedges)
   #type <- c(rep("simulated", (nrow(counts_data)-1)), rep("target", 1))
   #counts_data <- as.data.frame(cbind(counts_data, type))
   
  ggplot(counts_data)+
        geom_line(aes(x=time, y=main_edge_count*2/vertex_count))+
        geom_line(aes(x=time, y=nedges*2/n))+
        xlab("time step")+ylab("mean degree")
```

#### Degree distribution.
The simulated and target momentary degree distributions are below.
```{r, echo=FALSE, results='hide'}
sim_deg_dist_num <- degreedist(net)
sim_deg_dist <- sim_deg_dist_num/final.vcount
names(sim_deg_dist) <- c("0", "1", "2")
target_deg_dist_num <- deg_seq
target_deg_dist <- target_deg_dist_num/init.vcount
names(target_deg_dist) <- c("0", "1", "2")

main_deg_dist_data <- cbind(t(sim_deg_dist[1:3]), t(target_deg_dist))
main_deg_dist_data <- melt(main_deg_dist_data, id.vars=c("degree0, degree1, degree2"))
main_deg_dist_data <- cbind(main_deg_dist_data, c(rep("simulated", 3), rep("target", 3)))
colnames(main_deg_dist_data) <- c("ID", "degree", "value", "type")

ggplot(main_deg_dist_data, aes(x=degree, y=value, fill=type))+
  geom_bar(position="dodge", stat="identity")+
  ylab("Proportion of Agents")

```

#### Partnership duration.
```{r, echo=FALSE, results='hide'}
## (fast method from jonathan)

# looking for mean partnership durations

pt_data <- fread(paste0("../../Release/output/partnership_events_", sim_num,".csv")) 
#fread("../../Release/output/partnership_events_40.csv")
pt_data_main <- pt_data[network_type == 0]

# group by p1, p2
pt_data_main[,.(counts = .N),by=.(p1,p2),][,unique(counts)] # 1 2 4 3
pt_data_main[,.(counts = .N),by=.(p1,p2),][,.N,by=counts]

# counts     N
# 1:      1  5501
# 2:      2 21833
# 3:      4    10
# 4:      3     1

# so focus on just 2 is justified
    ## (old DONOTUSE: res <- pt_data_main[,.(tick, counts = .N),by=.(p1,p2),][counts == 2,abs(diff(tick))])

res <- pt_data_main[,.(tick, counts = .N),by=.(p1,p2),][counts == 2,.(dur = abs(diff(tick))), by=.(p1,p2)][,dur]

summary(res)

res_dt <- as.data.frame(res)
  qplot(res, geom="histogram", binwidth=100, 
       col=I("red"), 
       alpha=I(.2))

sum.res <- summary(res)

```

The target mean duration is `r duration`. In the simulated data above,
the interquartile range is `r as.numeric(sum.res[c(2,5)])`.  

### Casual Partnerships

#### Mean degree. 
The target casual mean degree is `r round(cas_n_edges*2/n, 2)` and a plot of the simulated mean
degree, over the course of the simulation, is below.

```{r, echo=FALSE, show='asis'}
   
  ggplot(counts_data)+
        geom_line(aes(x=time, y=casual_edge_count*2/vertex_count))+
        geom_line(aes(x=time, y=cas_n_edges*2/n))+
        xlab("time step")+ylab("mean degree")
```

#### Degree distribution.  
The simulated and target momentary degree distributions are below.

```{r, echo=FALSE, results='show'}
sim_deg_dist_num <- degreedist(cas_net)
sim_deg_dist <- sim_deg_dist_num/final.vcount
names(sim_deg_dist) <- c("0", "1", "2")
target_deg_dist_num <- cas_deg_seq
target_deg_dist <- target_deg_dist_num/init.vcount
names(target_deg_dist) <- c("0", "1", "2")

casual_deg_dist_data <- cbind(t(sim_deg_dist[1:3]), t(target_deg_dist[1:3]))
casual_deg_dist_data <- melt(casual_deg_dist_data, id.vars=c("degree0, degree1, degree2"))
casual_deg_dist_data <- cbind(casual_deg_dist_data, c(rep("simulated", 3), rep("target", 3)))
colnames(casual_deg_dist_data) <- c("ID", "degree", "value", "type")

ggplot(casual_deg_dist_data, aes(x=degree, y=value, fill=type))+
  geom_bar(position="dodge", stat="identity")+
  ylab("Proportion of Agents")

```

#### Partnership duration.
```{r, echo=TRUE, results='show'}
pt_data_casual <- pt_data[network_type == 1]

# group by p1, p2
pt_data_casual[,.(counts = .N),by=.(p1,p2),][,unique(counts)] # 1 2 4 3
pt_data_casual[,.(counts = .N),by=.(p1,p2),][,.N,by=counts]

# so focus on just 2 is justified
    ## (old DONOTUSE: res <- pt_data_casual[,.(tick, counts = .N),by=.(p1,p2),][counts == 2,abs(diff(tick))])

res <- pt_data_casual[,.(tick, counts = .N),by=.(p1,p2),][counts == 2,.(dur = abs(diff(tick))), by=.(p1,p2)][,dur]

summary(res)

res_dt <- as.data.frame(res)
  qplot(res, geom="histogram", binwidth=100, 
       col=I("red"), 
       alpha=I(.2))

sum.res <- summary(res)
```

The target mean duration is `r dur_cas`. In the simulated data above,
the interquartile range is `r as.numeric(sum.res[c(2,5)])`.  

### Overlap between main and casual partnerships.
```{r, echo=FALSE, results='show'}

overlap_counts <- as.data.frame(
  cbind(counts_data$time, counts_data$overlaps/counts_data$main_edge_count)
  )
overlap_counts <- rbind(overlap_counts, 
                        cbind(counts_data$time, counts_data$overlaps/counts_data$casual_edge_count))
denominator <- c(rep("main", nrow(overlap_counts)/2), 
                 rep("casual", nrow(overlap_counts)/2)
                 )
overlap_counts <- cbind(overlap_counts, denominator)
overlap_counts <- as.data.frame(overlap_counts)
colnames(overlap_counts) <- c("time", "overlap", "denominator")

ggplot(overlap_counts, aes(x=time, y=overlap, color=denominator))+
  geom_line()

```

### Total mean degree: main and casual partnerships combined.

```{r, echo=FALSE, results='show'}
qplot(data=counts_data, x=time, 
      y=((main_edge_count*2/vertex_count)+
           (casual_edge_count*2/vertex_count)), 
      main="Sim 36", ylab=("total mean degree"),
      geom="line")
```

## Testing and Diagnosis

```{r, echo=TRUE, results='show'}
p_inf_status <- person_data$infection.status
p_time_of_infection <- person_data$time.of.infection
p_art_status <- person_data$art.status
p_time_of_art_initiation <- person_data$time.of.art.initiation

table(p_inf_status)
table(p_art_status)

xtabs(~factor(p_inf_status, exclude=NULL) + 
        factor(p_art_status, exclude=NULL))

summary(p_time_of_art_initiation - p_time_of_infection)
neg_diff_art.init_inf <- neg.diff <- which(p_time_of_art_initiation - p_time_of_infection < 0)

p_inf_status[neg_diff_art.init_inf]
p_time_of_infection[neg_diff_art.init_inf]
p_time_of_art_initiation[neg_diff_art.init_inf]
summary(p_time_of_art_initiation)

```

## ART metrics

The ART initiation portion was simplified so that all infected's initate ART 1 year after seroconversion. At presetnt, the proportion of the infected on ART was

```{r, results='asis'}
  on.art <- which(net%v%"art.status" == 1)
  length(on.art)
  length(on.art)/length(infected)

```

`r round(length(on.art)/length(infected), 2)`. The total proportion of all people in the population who were on ART is `r round(length(on.art)/length(infected)*prev, 2)`%.

## Biomarkers



```{r, results='asis'}
   ## Selected upto n.to.follow infecteds at random
   n_to_follow <- 10
   infectees <- inf_event_data$infectee
   uniq_biom_pid <- unique(biom_data$p_id)

   infectee_uniq_biom_pid <- uniq_biom_pid[which(uniq_biom_pid %in% infectees)]
   new <- biom_data[which(biom_data$p_id == infectee_uniq_biom_pid[1]),]

   if (length(infectee_uniq_biom_pid > n_to_follow)){
     infectee_uniq_biom_pid <- infectee_uniq_biom_pid[1:n_to_follow]#first 10
     #infectee_uniq_biom_pid <- sample(infectee_uniq_biom_pid, 
                                      #n_to_follow, replace=FALSE)# random 10
   }
   
   for (i in 2:length(infectee_uniq_biom_pid)){
     new_entry <- biom_data[which(biom_data$p_id == infectee_uniq_biom_pid[i]),]
     new <- rbind(new, new_entry)
   }
   
   new$p_id <- as.character(new$p_id)

   par(mfrow=c(2,1))
   ggplot(new, aes(x=tick/365, y=viral_load, color=p_id))+
          geom_line()+xlab("year")
   ggplot(new, aes(x=tick/365, y=cd4_count, color=p_id))+
          geom_line()+xlab("year")

```

Example CD4 and viral load trajectories for `r n_to_follow` randomly selected, HIV-infected individuals are shown above. Individuals on ART are identifiable as those whose CD4 counts start to recover after decreasing initially; individuals who do not go on ART do experience a monotonic decrease.


# Conclusion
The features we have in the model appear to work well so far. We are in the process of adding more (as mentioned above), and continuing to monitor emergent trends in the behavior of key parameters and outcomes. 

Suggestions welcome! 
