
```{r, echo=FALSE}
rm(list=ls())

## libraries
suppressPackageStartupMessages(library(network))
suppressPackageStartupMessages(library(ergm))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(reshape2))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(data.table))
suppressPackageStartupMessages(library(plyr))
suppressPackageStartupMessages(library(ggthemes))
```


```{r, echo=FALSE}

loc <- "/project/khanna7/Projects/BARS/transmission_model/swift_proj/experiments/apr4_min_inf_zoomin/instance_9/output/"
 
ntimesteps <- 50000

## read data
   ## networks
   main_nets <- readRDS(paste0(loc, "/main_network_", ntimesteps, ".RDS"))
   cas_nets <- readRDS(paste0(loc, "/casual_network_", ntimesteps, ".RDS"))                    
                  
   ## other data
   biom_data <- fread(paste0(loc, "biomarker_log.csv"))
   counts_data <- fread(paste0(loc, "counts.csv"))                           
   inf_event_data <- fread(paste0(loc, "infection_events.csv"))                      
   pt_event_data <- fread(paste0(loc, "partnership_events.csv"))
   person_data <- fread(paste0(loc, "person_data.csv"))                                
   prep_events <- fread(paste0(loc, "prep_events.csv"))                           
   
   ## parameters
   source("../common/chi_params_nonderived.R")
   source("../common/chi_params_derived.R")
```

These datasets include the sexual network at the last time step, biomarker data with detailed trajectories, infection, death and partnership events, and counts of various quantities at each time step.

# Results
## Demography
We first compute the annual growth rate.  

```{r, echo=FALSE, results='hide'}
   
  final.vcount <- network.size(main_nets)
  init.vcount <- 5000
  
  annual.growth.rate <- (((final.vcount)/init.vcount)^(1/(10e3/365))-1)*100 #log scale ## CHICECK THIS
```


The age distribution at the end of the simulation is below.


```{r, results='show'}
# gg net%v%"age"
# plot(
#   qplot(age, geom="histogram", binwidth=5, 
#        col=I("red"), 
#        alpha=I(.2))
#   )
```

## Prevalence
    
```{r, echo=FALSE, results='hide'}

   alive_person_data <- person_data[time_of_death == -1] 
   head(alive_person_data); dim(alive_person_data)
   final_popsize <- nrow(alive_person_data)
   
   infected_data <- alive_person_data[infection_status == 1]                       
   n_infected_end <- nrow(infected_data)
   
   prev <- n_infected_end/final_popsize
```
   
  The overall prevalence at the end of the simulation was about `r round(prev, 1)`%.  HIV prevalence as a function of time is shown below.
  
  ```{r, echo=FALSE}
  counts_data <- counts_data[-1,]
     ggplot(counts_data, 
          aes(x=time/365, 
             y=(vertex_count-uninfected)/vertex_count))+
         geom_line()+
         scale_y_continuous(limits=c(0,1))+
         ylab("Prevalence")+
         xlab("Time (years)")+
         theme_minimal()
  ```
   
## Incidence  

The annual incidence rates over the course of the simulation are shown below.

   ```{r, results='hide', echo=FALSE}
  
      inc_per_timestep <- counts_data[, .((infected_via_transmission)/(uninfected - entries))] 
      summary(inc_per_timestep) #Compare computation time in vector and data.table form. 
      inc_per_timestep <- inc_per_timestep[["V1"]]
      #summ_inc_per_timestep <- lapply(inc_per_timestep, summary)
      
      inc_per_year <- split(inc_per_timestep, ceiling(seq_along(inc_per_timestep)/365))
      length(inc_per_year)
      list_mean_inc_per_year <- lapply(inc_per_year, mean)
      summ_inc_across_dur <- summary(unlist(list_mean_inc_per_year))*365*100

      list_mean_inc_per_year_dt <- as.data.frame(list_mean_inc_per_year)
      dt_mean_inc <- as.data.frame(t(unlist(list_mean_inc_per_year)))
      colnames(dt_mean_inc) <- "mean_daily_incidence"
      
      ggplot(as.data.frame(list_mean_inc_per_year_dt), 
             aes(y=mean_daily_incidence))+
             geom_line()
     
   ```
   
### Main Partnerships

#### Mean degree. 
The target main mean degree is `r round(nedges*2/n, 2)` and a plot of the simulated mean
degree, over the course of the simulations, is below.

```{r, echo=FALSE, show='hide'}
   #counts_data <- rbind(counts_data, nedges)
   #type <- c(rep("simulated", (nrow(counts_data)-1)), rep("target", 1))
   #counts_data <- as.data.frame(cbind(counts_data, type))
   
  ggplot(counts_data_df)+
        geom_line(aes(x=time/365, y=main_edge_count*2/vertex_count, col=sim_num))+
        geom_line(aes(x=time/365, y=nedges*2/n))+
        xlab("time step")+ylab("mean degree")+
        scale_y_continuous(limits=c(0.4, 0.5))+
        ylab("Mean degree (main)")+
        xlab("Time (years)")
  
```

#### Degree Distribution

```{r, echo=FALSE, show='hide'}
   sim_deg_dist_in_num <- lapply(main_nets, degreedist)
   sim_deg_dist_prop <- unlist(sim_deg_dist_in_num)/unlist(final.vcount)
   sim_deg_dist_prop_mat <- matrix(sim_deg_dist_prop, ncol=4, byrow=TRUE)

   sim_deg_dist_prop_mat_mean <- apply(sim_deg_dist_prop_mat, 2, mean)
   names(sim_deg_dist_prop_mat_mean) <- as.factor(0:3)
   
   target_deg_dist_in_num <- c(deg_seq, 0)
   target_deg_dist_prop <- target_deg_dist_in_num/init.vcount
   names(target_deg_dist_prop) <-  names(sim_deg_dist_prop_mat_mean) 
  
   main_deg_dist_data <- cbind(t(sim_deg_dist_prop_mat_mean), t(target_deg_dist_prop))
   main_deg_dist_data <- melt(main_deg_dist_data)
   main_deg_dist_data$type <- c(rep("mean_of_sim", 4), rep("target", 4))
   main_deg_dist_data_df <- as.data.frame(main_deg_dist_data)
   

   ggplot(main_deg_dist_data_df, aes(x=Var2, y=value, fill=type))+
    geom_bar(position="dodge", stat="identity")+
    ylab("Proportion of Agents")+
    xlab("Degree")

```

#### Partnership duration.
```{r, echo=FALSE, results='hide'}
## (fast method from jonathan)

# looking for mean partnership duration
pt_data_main <- lapply(pt_event_data, function (i)
                                      i[network_type == 0]
                       )
                       
# group by p1, p2
pt_main_counts <- lapply(pt_data_main, 
                        function (i)
                        i[,.(counts = .N),by=.(p1,p2),][,unique(counts)] # 1 2 4 3
                        )
pt_main_counts_n <- lapply(pt_data_main, 
                           function (i)
                           i[,.(counts = .N),by=.(p1,p2),][,.N,by=counts]
                           )
# counts     N
# 1:      1  5501
# 2:      2 21833
# 3:      4    10
# 4:      3     1

# so focus on just 2 is justified
    ## (old DONOTUSE: res <- pt_data_main[,.(tick, counts = .N),by=.(p1,p2),][counts == 2,abs(diff(tick))])

res_list <- lapply(pt_data_main, 
                   function (i)
                     i[,.(tick, counts = .N),by=.(p1,p2),][counts == 2,.(dur = abs(diff(tick))), by=.(p1,p2)][,dur]
                   )
summary(unlist(res_list))

res_dt <- as.data.frame(unlist(res_list))
  qplot(res_dt, geom="histogram", binwidth=100, 
       col=I("red"), 
       alpha=I(.2))

sum.res <- summary(unlist(res_list))

```

### Casual Partnerships

#### Mean degree. 
The target main mean degree is `r round(cas_n_edges*2/n, 2)` and a plot of the simulated mean
degree, over the course of the simulations, is below.

```{r, echo=FALSE, show='hide'}
   #counts_data <- rbind(counts_data, nedges)
   #type <- c(rep("simulated", (nrow(counts_data)-1)), rep("target", 1))
   #counts_data <- as.data.frame(cbind(counts_data, type))
   
  ggplot(counts_data_df)+
        geom_line(aes(x=time/365, y=casual_edge_count*2/vertex_count, col=sim_num))+
        geom_line(aes(x=time/365, y=cas_n_edges*2/n))+
        ylab("Mean degree (casual)")+
        xlab("Time (years)")
  
```


```{r, echo=FALSE, show='hide'}
   sim_cas_deg_dist_in_num <- lapply(cas_nets, degreedist)
   sim_cas_deg_dist_prop <- lapply(dir_nums, 
                               function (i) 
                               unlist(sim_cas_deg_dist_in_num[i])/unlist(final.vcount[i]))
   sim_cas_deg_dist_prop_df <- do.call("rbind.fill", 
                                        lapply(sim_cas_deg_dist_prop, as.data.frame))

    
   cas_n_deg <- unlist(lapply(sim_cas_deg_dist_prop, length))
   pholder <- list()
   list_cas_degs <- lapply(dir_nums, function (i) pholder[[i]] <- (0:(cas_n_deg[[i]] - 1)))
   sim_cas_deg_dist_prop_df$degree <- as.character(unlist(list_cas_degs))

   sim_cas_deg_mean <- aggregate(sim_cas_deg_dist_prop_df[,1], 
                                list(sim_cas_deg_dist_prop_df$degree), mean)
   colnames(sim_cas_deg_mean) <- c("degree", "mean")
   
   names(cas_deg_seq) <- as.character(0:(length(cas_deg_seq)-1))
   target_deg_dist_prop <- as.data.frame(cbind(names(cas_deg_seq),
                                         cas_deg_seq/init.vcount)
                                         )
   colnames(target_deg_dist_prop) <- colnames(sim_cas_deg_mean)

   casual_deg_dist_data <- rbind(sim_cas_deg_mean, target_deg_dist_prop)
   casual_deg_dist_data$type <- c(rep("mean_of_sim", nrow(sim_cas_deg_mean)), 
                                      rep("target", nrow(target_deg_dist_prop)))
   #casual_deg_dist_data_df <- as.data.frame(casual_deg_dist_data)
   
   class(casual_deg_dist_data$mean) <- "numeric"

   ggplot(casual_deg_dist_data, aes(x=degree, y=mean, fill=type))+
    geom_bar(position="dodge", stat="identity")+
    ylab("Proportion of Agents")+
    xlab("Degree")

```

#### PrEP outcomes

```{r, echo=FALSE, show='hide'}
head(prep_events)
dim(prep_events)

## How many (and what %) of uninfected individuals are on PrEP?
n_prep_use <- alive_person_data[infection_status == 0, .(.N), by=.(prep_status)] #n on PrEP
pct_prep_use <- n_prep_use[, . (percent = N/sum(N)*100)]

## Those who have started and stopped PrEP: How long did they stay on it? 
time_of_prep_use <- person_data[time_of_prep_initiation > 0 & time_of_prep_cessation > 0 & 
                                  prep_status == 0, 
                                (time_of_prep_use = time_of_prep_cessation - time_of_prep_initiation)
                                ]
summary(time_of_prep_use)/30

## Number of PrEP cycles
sort(colnames(prep_events))
head(prep_events)
dt_prep_events <- prep_events[, .(.N), by = p_id]
dt_prep_events <- dt_prep_events[N %% 2 == 0,] # consider only those who have started and completed a cycle
n_prep_cycles <- dt_prep_events[["N"]] %/% 2
hist(n_prep_cycles)

```
